NLO QCD+EW corrections to diphoton production in association with a vector boson

### 2017-12-05 results.tex
slightly modified in-text addition to 1) and moved it to footnote not to spoil the flow of the argument

In this section we present numerical results for the NLO QCD and NLO EW 
corrections to all three production processes of a diphoton pair in 
association with a third vector boson, a third photon, a $W$ or a $Z$ 
boson, at the LHC at a centre-of-mass energy of 13\,TeV. 
In case of an accompanying $W$ or a $Z$ boson, we consider the full 
off-shell leptonic final state, i.e.\ lepton-neutrino or lepton-pair 
production.
All results are obtained in the Standard Model using the complex-mass 
scheme \cite{Denner:2005fg} with the following input parameters
\begin{center}
  \begin{tabular}{rclrcl}
    $\alphazero$ &\shortequal& $1/137.03599976$  \qquad &&& \\
    $\Gmu$ &\shortequal& $1.1663787\times 10^{-5}\; \text{GeV}^2$ &&& \\
    $m_W$ &\shortequal& $80.385\; \text{GeV}$       & $\Gamma_W$ &\shortequal& $2.085\; \text{GeV}$ \\
    $m_Z$ &\shortequal& $91.1876\; \text{GeV}$      & $\Gamma_Z$ &\shortequal& $2.4952\; \text{GeV}$ \\
    $m_h$ &\shortequal& $125.0\; \text{GeV}$        & $\Gamma_h$ &\shortequal& $0$\\
    $m_t$ &\shortequal& $173.2\; \text{GeV}$        & $\Gamma_t$ &\shortequal& $0$\;.
  \end{tabular}
\end{center}
While we calculate triple photon production in the \alphazero-scheme, 
we use a mixed scheme for \aaw\ and \aaz production: at LO two 
powers of $\alpha$ are taken in the \alphazero-scheme, while 
one power of $\alpha$ is taken in the \Gmu-scheme. 
The additional power of $\alpha$ in the NLO EW correction 
is evaluated in \alphazero-scheme again. \revised{This choice is motivated by the fact
that the real emission contribution takes place at $q^2=0$ and in order to obtain
pole cancellation between the real emission contribution and the virtual corrections we
chose the same factors in both contributions. An alternative choice would be to use
the \Gmu-scheme throughout.}
The virtual amplitudes are renormalised correspondingly.
In all cases both the width of the top quark and the Higgs boson 
can safely be neglected as there are no diagrams containing either 
as $s$-channel propagators which can potentially go on-shell. 
All other lepton and parton masses and widths are set to zero, 
i.e.\ we are working in the five-flavour scheme.
We use the \textsc{CT14nlo} PDF set with $\alphas(m_Z)=0.118$ 
\cite{Dulat:2015mca}, interfaced through LHAPDF6 \cite{Buckley:2014ana}. 
The use of a QCD-only PDF is justified by the fact that, 
at LO, the photon induced corrections are either non-existent 
(\aaa, \aaw) or negligible (\aaz).
This finding will be detailed in Section \ref{sec:results:aaz}.

We define our central scales through
\begin{equation}
  \label{eq:murfdef}
  %\begin{split}
    \muR^0 = \muF^0 = %\left\{
    %\begin{array}{ll}
     % m_{\aaa} \qquad & \text{in \aaa\ production} \\[2mm]
      \tfrac{1}{2}\,H_T' \;. %& \text{in \aaw\ and \aaz\ production.}
    %\end{array}\right.
  %\end{split}
\end{equation}
In the case of the triple photon process $ H_\mathrm{T}' $ is just given by the scalar sum of all final state 
transverse momenta, for the two processes with the massive vector bosons it is defined as 
\begin{equation}
  \label{ew:defHT}
  \begin{split}
    H_\mathrm{T}' = E_\mathrm{T}^V + \sum_{\gamma,q,g} p_{\mathrm{T},i}
  \end{split}
\end{equation}
with $\left.E_\mathrm{T}^W\right.^2=(p_\ell+p_\nu)^2$ and 
$\left.E_\mathrm{T}^Z\right.^2=(p_{\ell^+}+p_{\ell^-})^2$ 
in full analogy to the case of the case of vector boson production 
in association with jets \cite{Berger:2009ep}.
As the Born process in each case has no $\muR$ dependence, we 
do not expect the choice of scale to have a significant influence 
on the size of the relative QCD and EW corrections.
We calculate the leading order cross section $\rd\sigma_\text{LO}(\muF)$, 
which only depends on the factorisation scale $\muF$, the 
NLO QCD differential correction factor $\deltaQCD(\muR,\muF)$, 
introducing the additional $\muR$-dependence, and the NLO EW 
differential correction factor $\deltaEW(\muF^0)$. 
To estimate the impact of yet-to-be-calculated higher-order 
corrections we vary the free scales $\muR$ and $\muF$ 
by the conventional factor of two around their central values 
$\muR^0$ and $\muF^0$, respectively. \revised{ $\muR$ and $\muF$ are
varied independently within a 7-point scale variation, i.e. the 
two cases where one scale is the inverse of the other scale is not present.
The uncertainty bands shown later in the various distributions are given
by the envelope of these 7 results. It should be noted that the QCD scale variation is
completely unrelated to the electroweak corrections. Nevertheless it is interesting
to compare the two as it is the general assumption the the LO QCD scale uncertainty
covers all higher order effect. Therefore it is interesting and important to investigate
whether this assumption is true. }
We do not vary the factorisation scale for the determination 
of $\deltaEW$ as the inherent, albeit normally phenomenologically 
irrelevant, stabilisation of the $\muF$-dependence at NLO EW 
is not reflected in our chosen PDF.
Hence, our NLO EW result exhibits the exact same $\muF$-dependence 
as the LO result.
We define 
\begin{equation}
  \label{eq:defnlo}
  \begin{split}
    \rd\sigma_\text{NLO QCD}
    \,=\;&\rd\sigma_\text{LO}\left(1+\deltaQCD\right)\\
    \rd\sigma_\text{NLO EW}
    \,=\;&\rd\sigma_\text{LO}\left(1+\deltaEW\right)\\
    \rd\sigma_\text{NLO \QCDpEW}
    \,=\;&\rd\sigma_\text{LO}\left(1+\deltaQCD+\deltaEW\right)\\
    \rd\sigma_\text{NLO \QCDtEW}
    \,=\;&\rd\sigma_\text{LO}\left(1+\deltaQCD\right)\left(1+\deltaEW\right)\;.
  \end{split}
\end{equation}
Therein, the difference between NLO \QCDpEW\ and NLO \QCDtEW, which 
is of relative $\order(\alphas\alpha)$, can serve as an indicator 
of the potential size of unknown corrections at that order.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.32\textwidth]{figs_aaa/pT_y1}
  \includegraphics[width=0.32\textwidth]{figs_aaa/pT_y2}
  \includegraphics[width=0.32\textwidth]{figs_aaa/pT_y3}
  \caption{
    Transverse momentum of the leading (left), subleading (centre) 
    and third leading (right) photon 
    in triple photon production at the LHC at 13\,TeV. 
    The distributions are shown at LO (blue), NLO QCD (green), 
    NLO EW (orange), NLO \QCDpEW\ (black) and NLO \QCDtEW\ (red) 
    including scale uncertainties. The top panel displays the 
    absolute predictions for both the case with \revised{(dashed)} and without \revised{(solid)} the 
    application of a jet veto of $p_\mathrm{T,jet}^\mathrm{veto}=\text{30\,GeV}$, 
    the latter ones are easily recognisable 
    due to their much reduced rate. The top ratio plot details 
    the relative corrections to the leading order cross section 
    without applying the jet veto, while the centre ratio plot 
    applies the jet veto. 
    The lower ratio highlights the size of the electroweak corrections.
    \revised{All ratios are taken with respect to the LO prediction 
    using out central scale choice.}
    \label{fig:aaa:pt}
  }
\end{figure}

@@

Finally, we consider diphoton production in association with 
a $Z$ boson in its leptonic decay channel. 
In practise we look at lepton pair production including 
all non-resonant diagrams.
The fiducial phase space of our analysis is defined as follows, again according 
to existing measurements \cite{Aad:2016sau}. 
First, we require the presence of exactly one lepton pair 
of opposite charge, dressed with all photons in a cone of 
size $R=0.1$, with $\pT>20\,\text{GeV}$ and $|\eta|<2.47$. 
The dressed lepton pair invariant mass must be larger than 40\,GeV. 
As in the case of \aaw production, we then require at least 
two identified photons among the remaining ones after dressing, 
using the procedure described in Sec.\ \ref{sec:results:aaa}, 
loosening the transverse momentum requirements to 
$\pT>15\,\text{GeV}$ for both the leading and the subleading 
identified photon. 
We further demand the angular separation of both photons 
to be $\Delta R(\gamma_1,\gamma_2)>0.4$ and each photon and 
each lepton to be $\Delta R(\ell,\gamma)>0.4$.
The inclusive cross sections and correction factors were 
detailed in Tab.\ \ref{tab:xsec}.

@@

In close analogy with the previous two processes considered, 
we start our discussion of the results with the transverse 
momenta of the leading and subleading photon and the charged 
lepton pair, cf.\ Figure \ref{fig:aaz:pt}. 
The first observation is, while the general behaviour of the 
spectra is the same as in \aaw production, the inclusive QCD 
corrections are much smaller in this case. 
They range between $\deltaQCD=0.3-0.4$ for the leading and 
subleading photon and rise to 0.7 for the lepton pair at 
500\,GeV. 
The restrictive jet veto employed for the previous process 
reduces the QCD corrections further. 
Beyond the negligible impact at very small transverse 
momenta, they turn negative early on, reaching $\deltaQCD=-0.2-0.4$ 
at 500\,GeV. 
The electroweak corrections are negative throughout, increasing 
from $\deltaEW=-4\%$ at low transverse momenta to $-20\%$ at 
500\,GeV.
They are consistently larger than the LO uncertainty estimate. 
As a consequence, the difference between the additive and 
multiplicative is comparably large, especially in the high 
\pT\ regions.

@@

Figure \ref{fig:aaz:myy} continues to display the invariant masses 
of the leading and subleading photon, and both the leading and 
subleading photon and the lepton pair. 
The familiar peak in $m_{\gamma_i\ell^+\ell^-}$ at the $Z$ boson mass 
is produced by 
configurations where the respective photon is emitted by the 
lepton-pair from a resonant $Z$ boson decay. 
Again, the QCD corrections are moderate, always staying below 
$\deltaQCD=0.4$ and of a similar shape as encountered already in 
\aaw\ production. 
The jet veto again reduces the QCD corrections and drives them negative 
in large portions of the observable range, reaching as much as 
$\deltaQCD=-0.5$ for the diphoton invariant mass. 
The electroweak corrections, on the other hand, are qualitatively very similar 
to the \aaw case: with the exception of the well known positive 
radiative corrections below the $Z$ peak, they are negative 
throughout, reaching $\deltaEW=-20\%$ at invariant masses of 
1\,TeV.
As in the case of the transverse momenta, due to the similar size 
of the QCD and electroweak corrections at large transverse 
momenta, the difference between their additive and the multiplicative 
combination is amplified.

>>>>

In this section we present numerical results for the NLO QCD and NLO EW 
corrections to all three production processes of a diphoton pair in 
association with a third vector boson, a third photon, a $W$ or a $Z$ 
boson, at the LHC at a centre-of-mass energy of 13\,TeV. 
In case of an accompanying $W$ or a $Z$ boson, we consider the full 
off-shell leptonic final state, i.e.\ lepton-neutrino or lepton-pair 
production.
All results are obtained in the Standard Model using the complex-mass 
scheme \cite{Denner:2005fg} with the following input parameters
\begin{center}
  \begin{tabular}{rclrcl}
    $\alphazero$ &\shortequal& $1/137.03599976$  \qquad &&& \\
    $\Gmu$ &\shortequal& $1.1663787\times 10^{-5}\; \text{GeV}^2$ &&& \\
    $m_W$ &\shortequal& $80.385\; \text{GeV}$       & $\Gamma_W$ &\shortequal& $2.085\; \text{GeV}$ \\
    $m_Z$ &\shortequal& $91.1876\; \text{GeV}$      & $\Gamma_Z$ &\shortequal& $2.4952\; \text{GeV}$ \\
    $m_h$ &\shortequal& $125.0\; \text{GeV}$        & $\Gamma_h$ &\shortequal& $0$\\
    $m_t$ &\shortequal& $173.2\; \text{GeV}$        & $\Gamma_t$ &\shortequal& $0$\;.
  \end{tabular}
\end{center}
While we calculate triple photon production in the \alphazero-scheme, 
we use a mixed scheme for \aaw\ and \aaz\ production: at LO two 
powers of $\alpha$ are taken in the \alphazero-scheme, while 
one power of $\alpha$ is taken in the \Gmu-scheme. 
The additional power of $\alpha$ in the NLO EW correction 
is evaluated in \alphazero-scheme again. \footnote{\revised{This choice is motivated by the fact
that the real emission contribution takes place at $q^2=0$ and in order to obtain
pole cancellation between the real emission contribution and the virtual corrections we
chose the same factors in both contributions. An alternative choice would be to use
the \Gmu-scheme for the NLO correction as well. This choice is tailored towards an 
improved description of the weak part of the electroweak correction. Although 
the difference of both schemes is beyond the accuracy of the present analysis, 
both schemes are easily converted from one into the other as 
$\deltaEW\propto\alpha$.
Thus, the electroweak corrections are a relative 3.7\% larger in this scheme.}}
The virtual amplitudes are renormalised correspondingly.
In all cases both the width of the top quark and the Higgs boson 
can safely be neglected as there are no diagrams containing either 
as $s$-channel propagators which can potentially go on-shell. 
All other lepton and parton masses and widths are set to zero, 
i.e.\ we are working in the five-flavour scheme.
We use the \textsc{CT14nlo} PDF set with $\alphas(m_Z)=0.118$ 
\cite{Dulat:2015mca}, interfaced through LHAPDF6 \cite{Buckley:2014ana}. 
The use of a QCD-only PDF is justified by the fact that, 
at LO, the photon induced corrections are either non-existent 
(\aaa, \aaw) or negligible (\aaz).
This finding will be detailed in Section \ref{sec:results:aaz}.

We define our central scales through
\begin{equation}
  \label{eq:murfdef}
  %\begin{split}
    \muR^0 = \muF^0 = %\left\{
    %\begin{array}{ll}
     % m_{\aaa} \qquad & \text{in \aaa\ production} \\[2mm]
      \tfrac{1}{2}\,H_T' \;. %& \text{in \aaw\ and \aaz\ production.}
    %\end{array}\right.
  %\end{split}
\end{equation}
In the case of the triple photon process $ H_\mathrm{T}' $ is just given by the scalar sum of all final state 
transverse momenta, for the two processes with the massive vector bosons it is defined as 
\begin{equation}
  \label{ew:defHT}
  \begin{split}
    H_\mathrm{T}' = E_\mathrm{T}^V + \sum_{\gamma,q,g} p_{\mathrm{T},i}
  \end{split}
\end{equation}
with $\left.E_\mathrm{T}^W\right.^2=(p_\ell+p_\nu)^2$ and 
$\left.E_\mathrm{T}^Z\right.^2=(p_{\ell^+}+p_{\ell^-})^2$ 
in full analogy to the case of the case of vector boson production 
in association with jets \cite{Berger:2009ep}.
As the Born process in each case has no $\muR$ dependence, we 
do not expect the choice of scale to have a significant influence 
on the size of the relative QCD and EW corrections.
We calculate the leading order cross section $\rd\sigma_\text{LO}(\muF)$, 
which only depends on the factorisation scale $\muF$, the 
NLO QCD differential correction factor $\deltaQCD(\muR,\muF)$, 
introducing the additional $\muR$-dependence, and the NLO EW 
differential correction factor $\deltaEW(\muF^0)$. 
To estimate the impact of yet-to-be-calculated higher-order 
corrections we vary the free scales $\muR$ and $\muF$ 
by the conventional factor of two around their central values 
$\muR^0$ and $\muF^0$, respectively. \revised{ $\muR$ and $\muF$ are
varied independently within a 7-point scale variation, i.e. the 
two cases where one scale is the inverse of the other scale is not present.
The uncertainty bands shown later in the various distributions are given
by the envelope of these 7 results. It should be noted that the QCD scale variation is
completely unrelated to the electroweak corrections. Nevertheless it is interesting
to compare the two as it is the general assumption the the LO QCD scale uncertainty
covers all higher order effects. Therefore it is interesting and important to investigate
whether this assumption is true. }
We do not vary the factorisation scale for the determination 
of $\deltaEW$ as the inherent, albeit normally phenomenologically 
irrelevant, stabilisation of the $\muF$-dependence at NLO EW 
is not reflected in our chosen PDF.
Hence, our NLO EW result exhibits the exact same $\muF$-dependence 
as the LO result.
We define 
\begin{equation}
  \label{eq:defnlo}
  \begin{split}
    \rd\sigma_\text{NLO QCD}
    \,=\;&\rd\sigma_\text{LO}\left(1+\deltaQCD\right)\\
    \rd\sigma_\text{NLO EW}
    \,=\;&\rd\sigma_\text{LO}\left(1+\deltaEW\right)\\
    \rd\sigma_\text{NLO \QCDpEW}
    \,=\;&\rd\sigma_\text{LO}\left(1+\deltaQCD+\deltaEW\right)\\
    \rd\sigma_\text{NLO \QCDtEW}
    \,=\;&\rd\sigma_\text{LO}\left(1+\deltaQCD\right)\left(1+\deltaEW\right)\;.
  \end{split}
\end{equation}
Therein, the difference between NLO \QCDpEW\ and NLO \QCDtEW, which 
is of relative $\order(\alphas\alpha)$, can serve as an indicator 
of the potential size of unknown corrections at that order.

@@

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.32\textwidth]{figs_aaa/pT_y1}
  \includegraphics[width=0.32\textwidth]{figs_aaa/pT_y2}
  \includegraphics[width=0.32\textwidth]{figs_aaa/pT_y3}
  \caption{
    Transverse momentum of the leading (left), subleading (centre) 
    and third leading (right) photon 
    in triple photon production at the LHC at 13\,TeV. 
    The distributions are shown at LO (blue), NLO QCD (green), 
    NLO EW (orange), NLO \QCDpEW\ (black) and NLO \QCDtEW\ (red) 
    including scale uncertainties. The top panel displays the 
    absolute predictions for both the case with \revised{(dashed)} and without \revised{(solid)} the 
    application of a jet veto of $p_\mathrm{T,jet}^\mathrm{veto}=\text{30\,GeV}$, 
    the latter ones are easily recognisable 
    due to their much reduced rate. The top ratio plot details 
    the relative corrections to the leading order cross section 
    without applying the jet veto, while the centre ratio plot 
    applies the jet veto. 
    The lower ratio highlights the size of the electroweak corrections.
    \revised{All ratios are taken with respect to the LO prediction 
    using the central scale choice.}
    \label{fig:aaa:pt}
  }
\end{figure}

@@

Finally, we consider diphoton production in association with 
a $Z$ boson in its leptonic decay channel. 
In practise we look at lepton pair production including 
all non-resonant diagrams.
The fiducial phase space of our analysis is defined as follows, again according 
to existing measurements \cite{Aad:2016sau}. 
First, we require the presence of exactly one lepton pair 
of opposite charge, dressed with all photons in a cone of 
size $R=0.1$, with $\pT>20\,\text{GeV}$ and $|\eta|<2.47$. 
The dressed lepton pair invariant mass must be larger than 40\,GeV. 
As in the case of \aaw\ production, we then require at least 
two identified photons among the remaining ones after dressing, 
using the procedure described in Sec.\ \ref{sec:results:aaa}, 
loosening the transverse momentum requirements to 
$\pT>15\,\text{GeV}$ for both the leading and the subleading 
identified photon. 
We further demand the angular separation of both photons 
to be $\Delta R(\gamma_1,\gamma_2)>0.4$ and each photon and 
each lepton to be $\Delta R(\ell,\gamma)>0.4$.
The inclusive cross sections and correction factors were 
detailed in Tab.\ \ref{tab:xsec}.

@@

In close analogy with the previous two processes considered, 
we start our discussion of the results with the transverse 
momenta of the leading and subleading photon and the charged 
lepton pair, cf.\ Figure \ref{fig:aaz:pt}. 
The first observation is, while the general behaviour of the 
spectra is the same as in \aaw\ production, the inclusive QCD 
corrections are much smaller in this case. 
They range between $\deltaQCD=0.3-0.4$ for the leading and 
subleading photon and rise to 0.7 for the lepton pair at 
500\,GeV. 
The restrictive jet veto employed for the previous process 
reduces the QCD corrections further. 
Beyond the negligible impact at very small transverse 
momenta, they turn negative early on, reaching $\deltaQCD=-0.2-0.4$ 
at 500\,GeV. 
The electroweak corrections are negative throughout, increasing 
from $\deltaEW=-4\%$ at low transverse momenta to $-20\%$ at 
500\,GeV.
They are consistently larger than the LO uncertainty estimate. 
As a consequence, the difference between the additive and 
multiplicative is comparably large, especially in the high 
\pT\ regions.

@@

Figure \ref{fig:aaz:myy} continues to display the invariant masses 
of the leading and subleading photon, and both the leading and 
subleading photon and the lepton pair. 
The familiar peak in $m_{\gamma_i\ell^+\ell^-}$ at the $Z$ boson mass 
is produced by 
configurations where the respective photon is emitted by the 
lepton-pair from a resonant $Z$ boson decay. 
Again, the QCD corrections are moderate, always staying below 
$\deltaQCD=0.4$ and of a similar shape as encountered already in 
\aaw\ production. 
The jet veto again reduces the QCD corrections and drives them negative 
in large portions of the observable range, reaching as much as 
$\deltaQCD=-0.5$ for the diphoton invariant mass. 
The electroweak corrections, on the other hand, are qualitatively very similar 
to the \aaw\ case: with the exception of the well known positive 
radiative corrections below the $Z$ peak, they are negative 
throughout, reaching $\deltaEW=-20\%$ at invariant masses of 
1\,TeV.
As in the case of the transverse momenta, due to the similar size 
of the QCD and electroweak corrections at large transverse 
momenta, the difference between their additive and the multiplicative 
combination is amplified.
